---
title: "reading_data_from_the_web"
author: "Kino Watanabe"
date: "2025-10-09"
output: html_document
---



```{r}
library(tidyverse)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
Import NSDOH data from the web
```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

drug_use_html
```
This is an "easy" case
```{r}
ndsuh_df = 
drug_use_html |>
  html_table() |> 
  nth(1) |> 
  slice(-1)
```
  * pulls every table in the url
  * `slice` takes off specific rows
      * “note” at the bottom of the table appears in every column in the first row
  
  * need to clean up "a, b, etc" and columns are encoding 2 different variables (age, year)
  
Slightly harder case

```{r}
swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```

Now pull out the elements I care about -- use selector gadget 
```{r}
title_vec = 
  swm_html |>
  html_elements(".ipc-title-link-wrapper .ipc-title__text") |>
  html_text()

metascore_vec = 
  swm_html |>
  html_elements(".metacritic-score-box") |>
  html_text()

runtime_vec = 
  swm_html |>
  html_elements(".dli-title-metadata-item:nth-child(2)") |>
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    score = metascore_vec,
    runtime = runtime_vec)
```

## APIs
Get data from APIs

NYC water consumption dataset
```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |> 
  content("parsed") 


nyc_water |> 
  ggplot(aes(x = year, y = nyc_consumption_million_gallons_per_day)) + 
    geom_point()
```

```{r}
brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) |> 
  content("parsed")
```
* ` query = list ("app_token" = xxxxxxxxxxxx))`


```{r}
poke = 
  GET("http://pokeapi.co/api/v2/pokemon/1") |>
  content()

poke[["name"]]
```
```{r}
poke[["height"]]
poke[["abilities"]]
```

```{r}

```

