---
title: "Strings and factors"
author: "Kino Watanabe"
date: "2025-10-20"
output: html_document
---

```{r}
library(rvest)
library(tidyverse)
library(p8105.datasets)
```

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## strings and regex
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

str_replace(string_vec, "jeff", "Jeff")

```

* str_detect() detects pattern with T/F
* str_replace() replace value

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")

```

* str_detect(string_vec, "^i think") detects "i think" only at the beginning of string
* str_detect(string_vec, "i think$") detects "i think" only at the end of string

* regular expressions are based on rules and text patterns

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")

```

* str_detect(string_vec,"[Pp]umpkin") --> can patch p or P if rest of string is "umpkin"


```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )
str_detect(string_vec, "^[0-9][a-zA-Z]")
```

* number first, followed IMMEDIATELY  by letter

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
str_detect(string_vec, "7\\.11")
```

* str_detect(string_vec, "7.11") --> "." is special char in regular expressions --> matches anything

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## factors
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

```

* can see integer structure of vec_sex

what happens if I relevel 
```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex

as.numeric(vec_sex)
```


## NSDUH -- strings
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |> #don't need p-values
  pivot_longer(
    -State, #pivot everything except for in the state column
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |> #separate age_year
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) #take out geo that's not states
```

## NSDUH -- factors

```{r}
data_marj |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> #start w char --> convert to factor --> reorder 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) #move x axis labels to 90 deg
```
## weather data

```{r}
data("weather_df")

weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
We could instead reorder name according to tmax values in each name using fct_reorder:

```{r}
weather_df |>
  mutate(name = forcats::fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
linear regression: the ordering determines the “reference” category, and is something that can be adjusted as needed.

```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  lm(tmax ~ name, data = _)

weather_df |>
  mutate(name = fct_reorder(name,tmax)) |> 
  lm(tmax ~ name, data = _)
```

```{r}
data("rest_inspec")

rest_inspec = 
    rest_inspec |> 
    rename(borough = boro)

rest_inspec |> 
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```


```{r}
rest_inspec =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), borough != "Missing") |> 
  mutate(borough = str_to_title(borough))
```
```{r}
rest_inspec |> 
  filter(str_detect(dba, "Pizza")) |> #match in str_detect is case-sensitive until we tell it not to be
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)

rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```


Might help to have things in a different order – maybe number of pizza places? Fortunately this can be done using fct_infreq.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar() 
```

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(borough = fct_infreq(borough)) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar() 
```

Suppose I want to rename a borough. I could try using str_replace.
This renamed the borough, but then converted the result back to a string 
```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    borough = fct_infreq(borough),
    borough = str_replace(borough, "Manhattan", "The City")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar()
```

```{r}
rest_inspec |> 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) |>
  mutate(
    borough = fct_infreq(borough),
    borough = fct_recode(borough, "The City" = "Manhattan")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar()
```

